---
title: "LDA ad QDA"
author: "I. Maslova"
date: "10/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Use the `Smarket`data like in Lab 6. Split the data into training and testing sets.

```{r}
library(ISLR2)
attach(Smarket)

train= Year<2005
test= !train

training_data=Smarket[train, -8]
testing_data=Smarket[test, -8]

testing_y=Direction[test]
```


## Linear Discriminant Analysis (LDA)

To run LDA in `R`, we will be using the `lda()` function which is part of a package called `MASS`. This package comes with the core distribution of `R`, so we just need to load it into `R`’s working directory before we use the `lda()` function.

```{r}
library(MASS)
lda.model<- lda(Direction~., data= training_data)
  
```

Next, we will assess the model, so again we will use `predict()` function on our testing data set.

```{r}
lda.pred<- predict(lda.model, testing_data)

names(lda.pred)
head(lda.pred)

lda.pred.Direction<-lda.pred$class
#don't need to assign true false like logistic regression, just assign class
```

The good news is that when using an LDA model in the `predict()` function, the output is the categories themselves(classes), unlike what happened when we used logistic regression model (the output was probabilities).

Alright, now it is time to assess the model. We create the confusion matrix and compute the misclassification error.

```{r}
table(lda.pred.Direction, testing_y)

# misclassification rate for LDA
mean(lda.pred.Direction!=testing_y)
```

## Quadratic Discriminant Analysis

Training and assessing a QDA model is very similar in syntax to training and assessing a QDA model. The only difference is in the function name `lda()`.

Let's build the model using the training data, predict the values for the testing set and assess the model performance.

```{r}
qda.model <- qda(Direction~., data = training_data)
qda.model

qda.pred = predict(qda.model, testing_data)

qda.pred.Direction=qda.pred$class
  
table(qda.pred.Direction, testing_y)

# compute the missclassification rate
mean(qda.pred.Direction!=testing_y)
```



